---
title: "Statistical Inference (Simulation Exercise)"
author: "EHarris"
date: "6/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Simulation Exercise Overview
This report will illustrate how well sample populations can provide a reliable picture or understanding of an entire population.  To do this, we will compare the sample means and sample variances for different sample sizes to the theoretical mean and variance.

Let's start with a simple comparison.  If we calculate the mean and variance of a large exponential sample population, like 10,000, we would expect this to be fairly close to the theoretical mean and variance.

**Large Sample Mean & Variance (10,000 observations)** 
```{r, large_sample_data, include = TRUE, echo = TRUE}
set.seed(100)
lrg.sample <- data.frame(rating = c(rexp(10000, 0.2)))

sample.mean <- round(mean(lrg.sample$rating), 3)   ## Population Mean  
sample.var <- round(var(lrg.sample$rating), 3)     ## Population Variance
sample.sd <- round(sd(lrg.sample$rating), 3)       ## Population Standard Deviation
```

As expected, the population mean of 10,000 observations is relatively close to the  
theoretical mean, `r sample.mean` is quite close to the theoretical mean of 5.  The same  
is true for variance.  The large sample variance of `r sample.var` is close to the  
theoretical variance of 25.

Let's look at a density distribution of the Random Exponsential values
```{r population_density_histogram, include = TRUE, echo = FALSE}
hist(lrg.sample$rating, prob = TRUE, breaks = 40, col = "red", xlab = "", ylim = c(0, .2),
     main = "Distribution of Random Exponentials (10k)", cex.main = 1.0)
abline(v = 5, col = "black", lwd = 2)
```

### Simulations
Instead of a single large sample, we will look at the average mean and average variance  
based on a varying number of simulations on sample of 40 Random Exponential values.  The  
number of simulations ranges from 1 to 1,000.  As the number of simulations increases  
from 1 to 1,000, it is expected that the average mean and average variance will more  
closely compare to the theoretical mean and variance.  Here is a list of attributes or  
parameters to define the sample populations:
1. Evaluate exponential distribution, using rexp(n, lambda), where  
+ n = 40 [sample of 40 exponentials]  
+ lambda = 0.2  
2. Mean, mu,ï­ of an exponential distribution is 1/lambda = 5 [1 / 0.2]  
3. Standard deviation (s) of an exponential distribution is 1/lambda = 5  
4. Variance, Var, of an exponential distribution is s-squared, 25

#### Sample Mean Simulations
We will run simulations where we calculate the mean of each sample of 40 random  
exponentials.

```{r sample.mean.forty, include = TRUE, echo = TRUE}
set.seed(100)
mlambda <- 0.2
msims <- 1000
n <- 40
mns = NULL
for (i in 1: msims) mns <- c(mns, mean(rexp(n,mlambda)))
```

##### Sample Mean versus Theoretical Mean
In the first simulation, we will calculate the mean for a single sample  
of 40 random exponentials.  The mean for this simulation may be considerably different  
from the theoretical mean.  However, in the final simulation, we will have 1,000 samples  
of 40 random exponentials.  It is anticipated the mean across each of the 1,000 samples  
will be more consistent with the theoretical mean.

Below is a comparison graph of the sample mean for each simulation to the theoretical  
mean.
```{r plot_simulation_mean_vs_theoretical, include = TRUE, echo = TRUE}
sim.mean <- cumsum(mns) / (1:msims)
plot(sim.mean, type = "l", lwd = 2, col = "black", ylim = c(4,5.5),
     xlab = "Number of Simulations",
     ylab = "Cumulative Mean",
     main = "Sample Mean vs. Theoretical Mean", cex.main = 1.0)
abline(h = 1/mlambda, lwd = 2, col = "red")         ## Line for Theoretical Mean
```

#### Sample Variance Simulations
Similar to the simulation for sample mean we are running simulations and calculating the  
average variance for simulations with 1 to 1,000 random samples of 40 exponential values.  

```{r sample.variance.forty, include = TRUE, echo = TRUE}
set.seed(100)
vlambda <- 0.2
vsims <- 1000
n <- 40
mns = NULL
for (i in 1: vsims) mns <- c(mns, var(rexp(n,vlambda)))
```

##### Sample Variance versus Theoretical Variance
In the first simulation, we will calculate the mean for a single sample  
of 40 random exponentials.  The mean for this simulation may be considerably different  
from the theoretical mean.  However, in the final simulation, we will have 1,000 samples  
of 40 random exponentials.  It is anticipated the mean across each of the 1,000 samples  
will be more consistent with the theoretical mean.

Below is a comparison graph of the sample mean for each simulation to the theoretical  
mean.
```{r plot_simulation_variance_vs_theoretical, include = TRUE, echo = TRUE}
sim.var <- cumsum(mns) / (1:vsims)
plot(sim.var, type = "l", lwd = 2, col = "black", ylim = c(10,30),
     xlab = "Number of Simulations",
     ylab = "Cumulative Variance",
     main = "Sample Variance vs. Theoretical Variance", cex.main = 1.0)
abline(h = (1/vlambda)^2, lwd = 2, col = "red")      ## Line for Theoretical Variance
```


