---
title: "Statistical Inference (Simulation Exercise)"
author: "EHarris"
date: "6/23/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Simulation Exercise Overview
This report will illustrate how well sample populations can provide a reliable picture or  
understanding of an entire population.  To do this, we will compare the sample means and  
sample variances for different sample sizes to the theoretical mean and variance.  We  
begin will the following information:

1. Evaluate exponential distribution, using rexp(n, lambda), where  
+ n = 40 [sample of 40 exponentials]  
+ lambda = 0.2  
2. Mean, mu, of an exponential distribution is 1/lambda = 5 [1 / 0.2]  
3. Standard deviation (s) of an exponential distribution is 1/lambda = 5  
4. Variance, Var, of an exponential distribution is s-squared, 25  

Let's start with a simple comparison.  If we calculate the mean and variance of a large  
exponential population, like 10,000, we would expect this to be fairly close to the  
theoretical mean and variance.

**Population Mean & Variance (10,000 observations)** 
```{r, population_stats, include = TRUE, echo = TRUE}
set.seed(123)
popdat <- data.frame(rating = c(rexp(10000, 0.2)))

popmean <- round(mean(popdat$rating), 3)   ## Population Mean  
popvar <- round(var(popdat$rating), 3)     ## Population Variance
```

As expected, the population mean of 10,000 observations is relatively close to the  
theoretical mean, `r popmean` is quite close to the theoretical mean of 5.  The same is  
true for variance.  A population variance of `r popvar` is close to the theoretical  
variance of 25.


