---
title: "Statistical Inference (Simulation Exercise)"
author: "EHarris"
date: "6/23/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Simulation Exercise Overview
This report will illustrate how well sample populations can provide a reliable picture or  
understanding of an entire population.  To do this, we will compare the sample means and  
sample variances for different sample sizes to the theoretical mean and variance.  We  
begin will the following information:

1. Evaluate exponential distribution, using rexp(n, lambda), where  
+ n = 40 [sample of 40 exponentials]  
+ lambda = 0.2  
2. Mean, mu, of an exponential distribution is 1/lambda = 5 [1 / 0.2]  
3. Standard deviation (s) of an exponential distribution is 1/lambda = 5  
4. Variance, Var, of an exponential distribution is s-squared, 25  

Let's start with a simple comparison.  If we calculate the mean and variance of a large  
exponential population, like 10,000, we would expect this to be fairly close to the  
theoretical mean and variance.

**Population Mean & Variance (10,000 observations)** 
```{r, large_sample_data, include = TRUE, echo = TRUE}
set.seed(123)
lrg.sample <- data.frame(rating = c(rexp(10000, 0.2)))

sample.mean <- round(mean(lrg.sample$rating), 3)   ## Population Mean  
sample.var <- round(var(lrg.sample$rating), 3)     ## Population Variance
sample.sd <- round(sd(lrg.sample$rating), 3)       ## Population Standard Deviation
```

As expected, the population mean of 10,000 observations is relatively close to the  
theoretical mean, `r sample.mean` is quite close to the theoretical mean of 5.  The same  
is true for variance.  The large sample variance of `r sample.var` is close to the  
theoretical variance of 25.

Let's look at a density distribution of the Random Exponsential values
```{r population_density_histogram, include = TRUE, echo = FALSE}
hist(lrg.sample$rating, prob = TRUE, breaks = 40, col = "red", xlab = "", ylim = c(0, .2),
     main = "Distribution of Random Exponentials (10k)", cex.main = 0.5)
abline(v = 5, col = "black", lwd = 2)




